---
title: "YPDC Session"
author: "Haseeb Raza"
date: "2024-07-08"
output:
  html_document:
    highlight: zenburn
    theme: cosmo
  pdf_document: default
---
# Importing iris data set
## The Iris dataset is a multivariate dataset
### The Iris dataset is a multivariate dataset introduced by the British biologist and statistician Ronald A. Fisher in his 1936 paper "The use of multiple measurements in taxonomic problems." It is often used as a beginner's dataset for machine learning and statistical classification techniques.

```{r}
head(iris)

```
```{r}
tail(iris)
```

```{r}
# Load the ggplot2 package
library(ggplot2)

 # Create a scatter plot of Petal.Length vs Petal.Width, colored by Species
ggplot(data = iris, aes(x = Petal.Length, y = Petal.Width, color = Species))+ 
  geom_point(size = 5, alpha = 0.5) +  # Add points with custom size and transparency
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") +  # Add linear model lines
  labs(
    title = "Scatter Plot of Iris Petal Dimensions",
    x = "Petal Length (cm)",
    y = "Petal Width (cm)"
  ) +
  theme_dark() +  # Use a minimal theme for a clean look
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))
```


```{r}
library(ggplot2)
library(plotly)
p1 <- # Load the ggplot2 package# Create a scatter plot of Petal.Length vs Petal.Width, colored by Species
ggplot(data = iris, aes(x = Petal.Length, y = Petal.Width, color = Species))+ 
  geom_point(size = 5, alpha = 0.5) +  # Add points with custom size and transparency
  geom_smooth(method = "lm", se = FALSE, linetype = "dashed") +  # Add linear model lines
  labs(
    title = "Scatter Plot of Iris Petal Dimensions",
    x = "Petal Length (cm)",
    y = "Petal Width (cm)"
  ) +
  theme_dark() +  # Use a minimal theme for a clean look
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

ggplotly(p1)
  
```

## here you can visit my linked in 

<a href="https://www.linkedin.com/in/haseeb-raza-344947297/?originalSubdomain=pk">Linkedin</a>

## using i frame tag for embedding other links




