---
title: "Statistics using R"
author: "Haseeb Raza"
date: "2024-07-11"
output:
  pdf_document: default
  html_document: 
    highlight: zenburn
    theme: journal
---
# T-test in statistics
## One sample t-test  & two sample t-test
### To check the normal distribution ;
to check the normality distribution we can perform Shapiro Wilk test and Kolmogrov Simirnov Test, 
or we can draw a histogram and qqnorm plot,

```{r}
df <- iris
shapiro.test(df$Sepal.Width)
```
so based upon the above result we can say that Sepal.Width in iris is normal.

```{r}
qqnorm(df$Sepal.Width)
```


```{r}
hist(df$Sepal.Width)
```


### To check the composition of data; 
we can perform Leven's Test to check the composition of variable
```{r}
library(car)
leveneTest(df$Sepal.Width ~ df$Species , data=df)
```

## One sample T-test on Sepal Width Data
```{r}
t.test(df$Sepal.Width , mu=5)
```

## Two sample T-test
There two types of two sample t-test 
- Unpaired T-test
- Paired T-test

### Unpaired T-test
```{r}
library(tidyverse)
t.test(df$Sepal.Width ~ df$Species=="setosa" |df$Species=="versicolor")
```

The above result indicates that these two variable has no significant difference.


```{r}
library(tidyverse)
t.test(df$Sepal.Width ~ df$Species=="versicolor" |df$Species=="virginica")
```

According to above the result the variables are significantly different from each other.

### Paired T-test 
For this test we have randomly created a data set having two thousand observations, with a standard deviation of ten.

```{r}
pre_treatment <- c(rnorm(2000, mean = 140, sd=10 ))
tibble(pre_treatment)
post_treatment <- c(rnorm(2000, mean = 170, sd=10))
tibble(post_treatment)
t.test(pre_treatment, post_treatment, paired = TRUE)

```

The result indicates that the averages of pre_treatment and post_treatment data differ significantly from each other.
